### 2016-06-30 responseToReviewers.Rmd

> _\textcolor{blue}{We reduced the introduction to limit the clinical motivation to a
TBI context and removed mention of multiple sclerosis to remove any assumption of
disease similarity (although the segmentation approaches can be similar).  We also
removed the paragraph dealing with publicly available software for neuroimaging analysis.
}_

2. Second, the manuscript has no results. All the results section is focused
on talking about feature ranking, which by the way is explained in the
Results section instead of the Methods section. The most important
results (the prediction accuracy of automated segmentation), are not
mentioned in the manuscript but are only plotted in Figure 7. My advice
is to put the focus of the results section on accuracy, i.e., give
precise dice values and perhaps add other measures (metric displacement,
Hausdorf distance, sensitivity, specificity, etc.). Also, the gain in
accuracy from stage 1 to stage 2 cannot be guessed by looking at a plot,
the authors must provide a statistical comparison. When investigating
accuracy, the error in lesion size should be reported, both in terms of
volume and in terms of individual lesions counts (i.e. out of 10 lesions
9 were found, 1 was missed, etc.). This information is crucial to
clinicians, if the authors foresee the use of the method in a clinical
setting, as it seems from the discussion. Beside adding results to the
manuscript, it is equally important to give some results in the abstract.
The abstract must be self-sufficient in reporting all the major points of
the paper.

> _\textcolor{blue}{We completely reorganized the results and discussion
section to reflect the new evaluation.  In addition, we added a synopsis
of these results to the abstract.}_

@@

3. Third, there is no information about the manual lesion tracing procedure.
Were lesion drawn on FLAIR or were other modalities used as well? If
FLAIR was used as reference it is not surprising that FLAIR emerges as
one of the top predictive modalities.

> _\textcolor{blue}{We added the following to the Methods section:}_

> \textcolor{blue}{The first author (J. R. S.) performed the manual WMH tracings for all 26  subjects.  J. R. S. is a radiologist certified by the American Board of Radiology, with a certificate of advanced qualification in vascular and interventional radiology, over 18 years of research experience in TBI, and 6 years of clinical imaging experience.  All multi-modal MR dicom image slices were converted to the nifti file format. All nifti image volumes for each subject were rigidly aligned to the T1 image of that subject using the ANTs software} [@Avants:2014aa] \textcolor{blue}{.  The normalized MRI volumes were then provided to J. R. S. who traced each lesion using the ITK-SNAP tool} [@Yushkevich:2006aa]
> \textcolor{blue}{ which has multi-image overlay capabilities for visualizing all modalities in all three canonical views.}

@@

\input{trainingData.tex}

\clearpage

4. Fourth, although the authors claim the method is available, I couldn't
find any link for the public. Note, it would be equally important if the
publication of the method online is associated with proper documentation
on how a new (naive) user can apply it.

> _\textcolor{blue}{Our methodological publication history demonstrates that we are
very supportive of open science and have released both code and data in conjunction
with these publications in public forums, such as github.  However, permission to
publicly release the data for this data has not yet been obtained and so we removed
the mention of public availability.  However, when we do get permission, it will
be posted to the github repository associated with this manuscript
(https://github.com/ntustison/WatchMeHyperventilate), which we mention in the text,
for future readers.
}_

@@

> _\textcolor{blue}{For this particular application, white matter hyperintensities
are the structure of interest.  We have re-emphasized this point by writing:
}_

> \textcolor{blue}{Crucial to these supervised segmentation approaches are the creation and selection of
``features'' as input (i.e., feature images constructed from the training data)
in conjunction with expertly identified structures of interest
(i.e., WMHs) for model construction.}

6. The methods section is very confusing. The section "Feature images for
WMH segmentation" starts with no explanation of what is Stage 1. The
logical flow would require template registration explained first, and all
feature computation performed later, and finally RF stages can be
explained with the available features.

> _\textcolor{blue}{We added Figure 1 which gives an overview of the basic workflow.
>   It can be seen below in response to item 2 of Reviewer 2.  We reference
>   this figure numerous times in the text as a visual explanatory aid.}_
> _\textcolor{blue}{We also added a paragraph at the beginning of the "Quantitative analysis" subsection
>   which gives an overview of the methods section including an description of Stage 1 and Stage2.}_
>
>   \textcolor{blue}{Figure 1 provides a graphical overview of the proposed workflow.  The major components
include offline generation of symmetric multimodal templates, the creation of
feature images from the training data which are then employed for statistical
prediction using a concatenated random forest framework.  This framework involves
the use of two random forest models where the outcome (i.e., the tissue
probability estimates) of the first RF model application,
which we denote as ``Stage1 '', is used as input (along with the original set of
feature images) to a refinement RF model segmentation which we denote as ``Stage 2''.
Once these offline steps
are performed, a new (i.e., unsegmented) subject can then be processed using the proposed pipeline.}
>
> _\textcolor{blue}{We also added a subsection explaining the symmetric multimodal template:}_
>
> \textcolor{blue}{Following} [@Tustison:2014ab] \textcolor{blue}{and} [@Tustison:2015aa], \textcolor{blue}{optimally derived templates
> serve for both brain tissue segmentation (for the derivation of tissue prior probability maps) and generation of asymmetry feature images.
> For this work we use the multi-modal data available from the public MMRR data set}
> [@landman2011]\textcolor{blue}{.  We used all 46 multi-modal acquisitions of that study to produce a multi-modal template according to the procedure described in} [@Avants:2010aa] \textcolor{blue}{which
> results in a mean (in terms of both shape and intensity) multivariate template representing the entire cohort.  Mid-canonical slices of the FLAIR, T1, and T2 components are illustrated in Figure 2.}
>
> _\textcolor{blue}{These changes complement the edits made in response to other review items.}_

@@

14. The discussion starts with an ``intent to further refine the proposed
paradigm,'' and proceeds by mentioning other work currently being
performed which looks promising. The opening of the discussion with such
statements truly make the manuscript sound like an incomplete effort that
is not ready for publication.

> _\textcolor{blue}{We appreciate the positive view of our current work.
The full sentence in question is ``Although only a
single expert was used to produce the manual labelings, our intent is to
further refine the proposed paradigm by crowdsourcing with feedback from
other experts who interact with both the data and methodology.''
In our opinion, simply mentioning specific details regarding future work
does not mean that the manuscript ``sounds like an incomplete effort.''
Rather, it is simply an admission, similar to much of what is reported
in the literature, that our research program is ongoing.}_

15. The discussion also spends several paragraphs on defining literature that
shows how WMHs are related to cognitive performance, or how WMHs can be
applied in diagnosing TBI. Yet, this study makes no effort to test these
efforts in these 24 patients. Consequently, those paragraphs sound
wishful and speculative. Given the methodological nature of this paper, a
more appropriate discussion would perhaps focus on the comparison with
other methods in terms of accuracy, similarities, differences, etc.

> _\textcolor{blue}{We removed the paragraphs in question and added a
comparison discussion.}_

### 2016-06-30 responseToReviewers.tex

\begin{quote}
\emph{\textcolor{blue}{We reduced the introduction to limit the clinical motivation to a
TBI context and removed mention of multiple sclerosis to remove any assumption of
disease similarity (although the segmentation approaches can be similar).  We also
removed the paragraph dealing with publicly available software for neuroimaging analysis.
}}
\end{quote}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{1}
\tightlist
\item
  Second, the manuscript has no results. All the results section is
  focused on talking about feature ranking, which by the way is
  explained in the Results section instead of the Methods section. The
  most important results (the prediction accuracy of automated
  segmentation), are not mentioned in the manuscript but are only
  plotted in Figure 7. My advice is to put the focus of the results
  section on accuracy, i.e., give precise dice values and perhaps add
  other measures (metric displacement, Hausdorf distance, sensitivity,
  specificity, etc.). Also, the gain in accuracy from stage 1 to stage 2
  cannot be guessed by looking at a plot, the authors must provide a
  statistical comparison. When investigating accuracy, the error in
  lesion size should be reported, both in terms of volume and in terms
  of individual lesions counts (i.e.~out of 10 lesions 9 were found, 1
  was missed, etc.). This information is crucial to clinicians, if the
  authors foresee the use of the method in a clinical setting, as it
  seems from the discussion. Beside adding results to the manuscript, it
  is equally important to give some results in the abstract. The
  abstract must be self-sufficient in reporting all the major points of
  the paper.
\end{enumerate}

\begin{quote}
\emph{\textcolor{blue}{We completely reorganized the results and discussion
section to reflect the new evaluation.  In addition, we added a synopsis
of these results to the abstract.}}
\end{quote}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{2}
\tightlist
\item
  Third, there is no information about the manual lesion tracing
  procedure. Were lesion drawn on FLAIR or were other modalities used as
  well? If FLAIR was used as reference it is not surprising that FLAIR
  emerges as one of the top predictive modalities.
\end{enumerate}

\begin{quote}
\emph{\textcolor{blue}{We added the following to the Methods section:}}
\end{quote}

\begin{quote}
\textcolor{blue}{The first author (J. R. S.) performed the manual WMH tracings for all 26  subjects.  J. R. S. is a radiologist certified by the American Board of Radiology, with a certificate of advanced qualification in vascular and interventional radiology, over 18 years of research experience in TBI, and 6 years of clinical imaging experience.  All multi-modal MR dicom image slices were converted to the nifti file format. All nifti image volumes for each subject were rigidly aligned to the T1 image of that subject using the ANTs software}\textsuperscript{1}
\textcolor{blue}{.  The normalized MRI volumes were then provided to J. R. S. who traced each lesion using the ITK-SNAP tool}\textsuperscript{2}
\textcolor{blue}{ which has multi-image overlay capabilities for visualizing all modalities in all three canonical views.}
\end{quote}

@@

\input{trainingData.tex}

\clearpage

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{3}
\tightlist
\item
  Fourth, although the authors claim the method is available, I couldn't
  find any link for the public. Note, it would be equally important if
  the publication of the method online is associated with proper
  documentation on how a new (naive) user can apply it.
\end{enumerate}

\begin{quote}
\emph{\textcolor{blue}{Our methodological publication history demonstrates that we are
very supportive of open science and have released both code and data in conjunction
with these publications in public forums, such as github.  However, permission to
publicly release the data for this data has not yet been obtained and so we removed
the mention of public availability.  However, when we do get permission, it will
be posted to the github repository associated with this manuscript
(https://github.com/ntustison/WatchMeHyperventilate), which we mention in the text,
for future readers.
}}
\end{quote}

@@

\begin{quote}
\textcolor{blue}{Crucial to these supervised segmentation approaches are the creation and selection of
``features'' as input (i.e., feature images constructed from the training data)
in conjunction with expertly identified structures of interest
(i.e., WMHs) for model construction.}
\end{quote}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{5}
\tightlist
\item
  The methods section is very confusing. The section ``Feature images
  for WMH segmentation'' starts with no explanation of what is Stage 1.
  The logical flow would require template registration explained first,
  and all feature computation performed later, and finally RF stages can
  be explained with the available features.
\end{enumerate}

\begin{quote}
\emph{\textcolor{blue}{We added Figure 1 which gives an overview of the basic workflow.
  It can be seen below in response to item 2 of Reviewer 2.  We reference
  this figure numerous times in the text as a visual explanatory aid.}}
\emph{\textcolor{blue}{We also added a paragraph at the beginning of the "Quantitative analysis" subsection
  which gives an overview of the methods section including an description of Stage 1 and Stage2.}}

\textcolor{blue}{Figure 1 provides a graphical overview of the proposed workflow.  The major components
include offline generation of symmetric multimodal templates, the creation of
feature images from the training data which are then employed for statistical
prediction using a concatenated random forest framework.  This framework involves
the use of two random forest models where the outcome (i.e., the tissue
probability estimates) of the first RF model application,
which we denote as ``Stage1 '', is used as input (along with the original set of
feature images) to a refinement RF model segmentation which we denote as ``Stage 2''.
Once these offline steps
are performed, a new (i.e., unsegmented) subject can then be processed using the proposed pipeline.}

@@

\emph{\textcolor{blue}{We also added a subsection explaining the symmetric multimodal template:}}

\textcolor{blue}{Following}\textsuperscript{3}
\textcolor{blue}{and}\textsuperscript{4},
\textcolor{blue}{optimally derived templates
serve for both brain tissue segmentation (for the derivation of tissue prior probability maps) and generation of asymmetry feature images.
For this work we use the multi-modal data available from the public MMRR data set}\textsuperscript{5}\textcolor{blue}{.  We used all 46 multi-modal acquisitions of that study to produce a multi-modal template according to the procedure described in}\textsuperscript{6}
\textcolor{blue}{which
results in a mean (in terms of both shape and intensity) multivariate template representing the entire cohort.  Mid-canonical slices of the FLAIR, T1, and T2 components are illustrated in Figure 2.}

@@

\emph{\textcolor{blue}{These changes complement the edits made in response to other review items.}}
\end{quote}

\begin{quote}
The pipeline also seem to be repetitive, the N4 correction is mentioned
in ln. 26 and then again in ln. 43, which sounds like it was performed
twice. Tissue segmentation is mentioned in ln. 39, then it is stated is
a Bayesian method (ln 45), and then it is referred to as Atropos in the
next page.
\end{quote}

\begin{quote}
\emph{\textcolor{blue}{The two mentions of N4 are due to the fact that they are associated
with two distinct pipelines:  1) preparation of the images for extracting feature values
and 2) brain extration and $n$-tissue segmentation.  We rewrote the section in question
to avoid any confusing redundancy:}}
\end{quote}

\begin{quote}
\emph{\textcolor{blue}{The T1 image is then processed via the ANTs brain
extraction and normal tissue segmentation pipelines}}\textsuperscript{3}
\emph{\textcolor{blue}{.  The result is a mask delineating the brain
parenchyma and probabilistic estimates of the CSF, gray matter, white matter,
deep gray matter, brain stem, and cerebellum}}\textsuperscript{7}
\emph{\textcolor{blue}{.  These provide the expertly annotated labels for the
first six tissue labels given above.  Tissue prior probability maps for segmentation
are from multi-model optimal symmetric shape/intensity templates}}\textsuperscript{4}
\emph{\textcolor{blue}{created from the public MMRR data set}}\textsuperscript{5}
\emph{\textcolor{blue}{(cf Figure 3).}}
\end{quote}

@@

\begin{quote}
\emph{\textcolor{blue}{Done.  We removed the mathematical formulas.}}
\end{quote}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{6}
\tightlist
\item
  Random forests are explained in a way that can be understood only by
  readers who already know how they work; a new reader (especially a
  clinician) would not be able to understand the method. For example,
  data pushed onto a tree is does not mean much, unless the reader knows
  what is an identification tree.
\end{enumerate}

\begin{quote}
\emph{\textcolor{blue}{
We revised the explanation of random forests as follows:}}
\end{quote}

\begin{quote}
\textcolor{blue}{The basic component of
the random forest paradigm is the ``decision tree'' often represented by a flowchart or
graph where internal nodes represent ``tests'', or decisions, and the edges represent the outcome of those tests.  The final, or end, nodes represent the various classsifications produced by traversal through the decision tree.  For the proposed application, individual
voxels (and their corresponding feature values) are introduced at the root of a particular
decision tree and traverse the edges and internal nodes ultimately ending up at one of
the classification nodes according to the tests at each internal node.  A single random
forest model will consist of many such trees (often refered to as an ``ensemble'').}
\textcolor{blue}{Although decision trees had been extensively studied, the success of employing collections of such weak learners for boosting machine learning performance
(e.g., AdaBoost)}\textsuperscript{8,9}
\textcolor{blue}{influenced the similarly sytled conglomeration of decision trees into ``forests'' with randomized node optimization}\textsuperscript{10,11}.
\textcolor{blue}{Finally, Breiman}\textsuperscript{12}
\textcolor{blue}{improved accuracy by random sampling of training data (i.e., ``bagging'') resulting
in the current random forest technique applied here.}
\textcolor{blue}{As voxels and their feature values are ``pushed'' through each decision tree in the forest,  votes for each label are accumulated and converted to probability values for all classification possiblilities at
each voxel location.}
\end{quote}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{7}
\tightlist
\item
  The relationship between WMH and GOS-E is mentioned in pg. 2 but the
  reader is not informed what is GOS-E. A similar issue is met few lines
  after, where the ``outcome'' is mentioned, but there is no information
  what is referred to as outcome.
\end{enumerate}

\begin{quote}
\emph{\textcolor{blue}{The sentence in question reads:}}
\end{quote}

\begin{quote}
\textcolor{blue}{Further, volume of FLAIR lesions within the corpus
callosum, brainstem, and thalamus in patients with severe TBI correlates
with Glasgow Outcome-Extended (GOS-E) scores.}\textsuperscript{13}
\end{quote}

@@

\begin{quote}
\emph{\textcolor{blue}{Although the full phrase corresponding to the acronym
in question is provided prior to the acronym in the text (along with an
explanatory reference), we have added the following clarification:}}
\end{quote}

\begin{quote}
\textcolor{blue}{Further, volume of FLAIR lesions within the corpus
callosum, brainstem, and thalamus in patients with severe TBI correlates
with Glasgow Outcome-Extended (GOS-E) scores}\textsuperscript{13}
\textcolor{blue}{---a numeric groupwise assessment used to
classify ``outcome'' in TBI patients where ``outcome'' refers to the
spectrum of possible prognoses from death to disability to recovery.}
\end{quote}

@@

\begin{quote}
\emph{\textcolor{blue}{This was a grammatical mistake, not a question of
ethics.  It has been fixed: ``The feature images
were derived from MR acquisitions of 26 subjects\ldots''}}
\end{quote}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{10}
\tightlist
\item
  It is unclear to me the meaning of the sentence: ``Since the inverse
  transform is also derived as part of the registration process, we can
  warp the voxel index locations back to the space of the individual
  subject which motivates similar work by others {[}43{]}.''
\end{enumerate}

\begin{quote}
\emph{\textcolor{blue}{The sentences were rewritten as follows:}}
\end{quote}

\begin{quote}
\textcolor{blue}{Similar feature images were used in}\textsuperscript{14}
\textcolor{blue}{although, unlike the proposed framework, this previous work lacks
normalization to the standard
coordinate system provided by the template to dramatically improve spatial specificity
across all subjects.  To generate these images, the T1 image of each subject is
registered to the T1 template component using a B-spline variant}\textsuperscript{15}
\textcolor{blue}{of the well-known ANTs Symmetric Normalization (SyN) algorithm}\textsuperscript{16}.
\textcolor{blue}{Using the derived transforms, the template coordinate images are warped back to the space of the individual subject.}
\end{quote}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{13}
\tightlist
\item
  The discussion starts with an ``intent to further refine the proposed
  paradigm,'' and proceeds by mentioning other work currently being
  performed which looks promising. The opening of the discussion with
  such statements truly make the manuscript sound like an incomplete
  effort that is not ready for publication.
\end{enumerate}

\begin{quote}
\emph{\textcolor{blue}{We appreciate the positive view of our current work.
The full sentence in question is ``Although only a
single expert was used to produce the manual labelings, our intent is to
further refine the proposed paradigm by crowdsourcing with feedback from
other experts who interact with both the data and methodology.''
In our opinion, simply mentioning specific details regarding future work
does not mean that the manuscript ``sounds like an incomplete effort.''
Rather, it is simply an admission, similar to much of what is reported
in the literature, that our research program is ongoing.}}
\end{quote}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{14}
\tightlist
\item
  The discussion also spends several paragraphs on defining literature
  that shows how WMHs are related to cognitive performance, or how WMHs
  can be applied in diagnosing TBI. Yet, this study makes no effort to
  test these efforts in these 24 patients. Consequently, those
  paragraphs sound wishful and speculative. Given the methodological
  nature of this paper, a more appropriate discussion would perhaps
  focus on the comparison with other methods in terms of accuracy,
  similarities, differences, etc.
\end{enumerate}

\begin{quote}
\emph{\textcolor{blue}{We removed the paragraphs in question and added a
comparison discussion.}}
\end{quote}

@@

\begin{figure}[htbp]
\centering
\includegraphics{Figures/wmhPipeline.png}
\caption{\textcolor{blue}{Workflow illustration for the proposed pipeline.  Processing of the multi-modal
input MRI for a single subject, using the multi-modal symmetric template, results in
the generation of the feature images.  These feature images are used as input to the
Stage 1 RF model producing the initial RF probability map estimates.  The Stage 1
voting maps, the original feature images, and the Stage 2 RF model result in the
final voting maps which includes the WMH probability estimate.  Note that the RF models
are constructed once from a set of training data which are processed using the
same feature-construction pipeline as the single-subject input MRI.}}
\end{figure}

\clearpage

\subsection*{References}\label{references}
\addcontentsline{toc}{subsection}{References}

\hypertarget{refs}{}
\hypertarget{ref-Avants:2014aa}{}
1. Avants BB, Tustison NJ, Stauffer M, Song G, Wu B, Gee JC. The Insight
ToolKit image registration framework. Front Neuroinform. 2014;8:44.

\hypertarget{ref-Yushkevich:2006aa}{}
2. Yushkevich PA, Piven J, Hazlett HC, Smith RG, Ho S, Gee JC, Gerig G.
User-guided 3D active contour segmentation of anatomical structures:
Significantly improved efficiency and reliability. Neuroimage.
2006;31(3):1116--28.

\hypertarget{ref-Tustison:2014ab}{}
3. Tustison NJ, Cook PA, Klein A, Song G, Das SR, Duda JT, Kandel BM,
Strien N van, Stone JR, Gee JC, et al. Large-scale evaluation of aNTs
and freeSurfer cortical thickness measurements. Neuroimage.
2014;99:166--79.

\hypertarget{ref-Tustison:2015aa}{}
4. Tustison NJ, Shrinidhi KL, Wintermark M, Durst CR, Kandel BM, Gee JC,
Grossman MC, Avants BB. Optimal symmetric multimodal templates and
concatenated random forests for supervised brain tumor segmentation
(simplified) with aNTsR. Neuroinformatics. 2015;13(2):209--25.

\hypertarget{ref-landman2011}{}
5. Landman BA, Huang AJ, Gifford A, Vikram DS, Lim IAL, Farrell JAD,
Bogovic JA, Hua J, Chen M, Jarso S, et al. Multi-parametric neuroimaging
reproducibility: A 3-T resource study. Neuroimage. 2011;54(4):2854--66.

\hypertarget{ref-Avants:2010aa}{}
6. Avants BB, Yushkevich P, Pluta J, Minkoff D, Korczykowski M, Detre J,
Gee JC. The optimal template effect in hippocampus studies of diseased
populations. Neuroimage. 2010;49(3):2457--66.

\hypertarget{ref-Avants:2011aa}{}
7. Avants BB, Tustison NJ, Wu J, Cook PA, Gee JC. An open source
multivariate framework for \(n\)-tissue segmentation with evaluation on
public data. Neuroinformatics. 2011;9(4):381--400.

\hypertarget{ref-schapire1990}{}
8. Schapire R. The strength of weak learnability. Machine Learning.
1990;5:197--227.

\hypertarget{ref-freund1997}{}
9. Freund Y, Schapire R. A decision-theoretic generalization of on-line
learning and an application to boosting. Journal of Computer and System
Sciences. 1997;55:119--139.

\hypertarget{ref-ho1995}{}
10. Ho TK. Random decision forests. In: Document analysis and
recognition, 1995., proceedings of the third international conference
on. Vol. 1. 1995. pp. 278--282 vol.1.

\hypertarget{ref-amit1997}{}
11. Amit Y, Geman D. Shape quantization and recognition with randomized
trees. Neural Computation. 1997;9:1545--1588.

\hypertarget{ref-breiman2001}{}
12. Breiman L. Random forests. In: Machine learning. 2001. pp. 5--32.

\hypertarget{ref-Moen:2014aa}{}
13. Moen KG, Brezova V, Skandsen T, Håberg AK, Folvik M, Vik A.
Traumatic axonal injury: The prognostic value of lesion load in corpus
callosum, brain stem, and thalamus in different magnetic resonance
imaging sequences. J Neurotrauma. 2014;31(17):1486--96.

\hypertarget{ref-Anbeek:2004aa}{}
14. Anbeek P, Vincken KL, Osch MJP van, Bisschops RHC, Grond J van der.
Probabilistic segmentation of white matter lesions in mR imaging.
Neuroimage. 2004;21(3):1037--44.

### 2016-06-30 responseToReviewers.tex

\hypertarget{ref-Tustison:2013ac}{}
15. Tustison NJ, Avants BB. Explicit B-spline regularization in
diffeomorphic image registration. Front Neuroinform. 2013;7:39.

\hypertarget{ref-Avants:2011ab}{}
16. Avants BB, Tustison NJ, Song G, Cook PA, Klein A, Gee JC. A
reproducible evaluation of ANTs similarity metric performance in brain
image registration. Neuroimage. 2011;54(3):2033--44.

### 2016-06-30 abstract.Rmd
ENH:  More revisions.

White matter hyperintensities (WMHs) are foci of abnormal signal
intensity in white matter regions seen with magnetic resonance imaging
(MRI). WMHs are associated with normal aging and have shown prognostic
value in neurological conditions such as traumatic brain injury (TBI).
The impracticality of manually quantifying these lesions limits their
clinical utility and motivates the utilization of machine learning
techniques for automated segmentation workflows.  Herein, we develop a
concatenated random forest framework with image features for
segmenting WMHs in a TBI cohort. The framework is built upon the
Advanced Normalization Tools (ANTs)  and ANTsR toolkits.  MR
(3D FLAIR, T2-, and T1-weighted) images from 24 service members and
veterans scanned in the Chronic Effects of Neurotrauma Consortium’s
(CENC) observational study were acquired. Manual annotations were
employed for both training and evaluation using a leave-one-out
strategy.  Lesion load and overlap evaluative comparisons are
complimented by feature rankings which showcase the utility of the
concatenated approach.  Our findings suggest supervised learning methods
may be applied to quantify WMHs on routine brain imaging. Paired with
correlative outcome data, supervised learning methods may allow for
identification of imaging features predictive of diagnosis and prognosis
in individual TBI patients.

>>>>

White matter hyperintensities (WMHs) are foci of abnormal signal
intensity in white matter regions seen with magnetic resonance imaging
(MRI). WMHs are associated with normal aging and have shown prognostic
value in neurological conditions such as traumatic brain injury (TBI).
The impracticality of manually quantifying these lesions limits their
clinical utility and motivates the utilization of machine learning
techniques for automated segmentation workflows.  Herein, we develop a
concatenated random forest framework with image features for
segmenting WMHs in a TBI cohort. The framework is built upon the
Advanced Normalization Tools (ANTs)  and ANTsR toolkits.  MR
(3D FLAIR, T2-, and T1-weighted) images from 24 service members and
veterans scanned in the Chronic Effects of Neurotrauma Consortium’s
(CENC) observational study were acquired. Manual annotations were
employed for both training and evaluation using a leave-one-out
strategy.  \textcolor{blue}{Performance measures include sensitivity,
positive predictive value, $F_1$ score, and relative volume difference.
Final average results were: sensitivity = 0.68 $\pm$ 0.38, positive
predictive value = 0.51 $\pm$ 0.40, $F_1$ = 0.52 $\pm$ 0.36,
relative volume difference} = 43 $\pm$ 26\%.  In addition, three
lesion size ranges are selected to illustrate the variation in
performance with lesion size.}  Paired with
correlative outcome data, supervised learning methods may allow for
identification of imaging features predictive of diagnosis and prognosis
in individual TBI patients.

### 2016-06-30 discussion.Rmd
ENH:  More revisions.

@@

>>>>

\textcolor{blue}{
In evaluating the segmentation results, there are limited methodologies described in the
literature for automatic segmentation of WMHs despite the numerous clinical studies}
[@Debette:2010aa;@Kloppenborg:2014aa] \textcolor{blue}{exploring the connections between WMHs and TBI.
However, we can extrapolate from similar application areas where such methodologies have a
much longer history of development such as white matter lesion segmentation in
multiple sclerosis )MS).  In a recent review [@Garcia-Lorenzo:2013aa], 47 total papers
were found during a literature search and evaluated which concern automated segmentation of white
matter lesions in MS.  The most widely-used data for algorithmic evaluation consists
of two cohorts from two different sites used in
the MS Lesion segmentation challenge associated with the international MICCAI 2008
conference.  For comparison, this challenge training data set
has a mean lesion load of 204 ($\pm$ 752) mm$^3$ per lesion (compared with
our mean lesion load of 33.8 ($\pm$ 55.4) voxels $\times$ 1.2 mm$^3$ per voxel =
40.56 ($\pm$ 66.48) mm$^3$ per lesion) and the resolution is
almost twice what is used in this study (i.e., 0.5 $\times$ 0.5 $\times$ 0.5).
Note that our relative volume difference numbers are comparable to the
relative volume difference {\em between raters} of 68 percent (cf Figure 5 (d))}
[@vanginneken:2007aa].

Regarding the feature rankings, it is interesting to note some of the other top performing features for Stage 1.
The contralateral difference FLAIR image is  highly discriminative over the set of
evaluation random forest models (see Figure 8).  This accords with the known clinical relevance of
FLAIR images for identifying white matter hyperintensities and the fact that such
pathology does not typically manifest symmetrically in both hemispheres.  Interestingly, the
posterior maps for the deep gray matter are extremely important for accurate white matter
hyperintensity segmentation.  Perhaps the spatial specification of deep
gray matter aids in the removal of false positives.  Inspection of the bottom of the
plots demonstrates the lack of discriminating features associated with the
T1 image which is also well-known in the clinical literature.

![(a) FLAIR image slice illustrating WMHs which have been manually delineated.
The region around the WMHs is enlarged (b) in the original FLAIR and the
(c) contralateral FLAIR difference image.](Figures/FLAIRcontralaleteralWithLesionsAlternate.png)

As described earlier, for Stage 2, we used the output random forest voting maps from Stage
1 as both features themselves and as priors for input to a Bayesian-based segmentation with
an additional MRF spatial prior.  In Figure 7, the voting maps are labeled as
"_RFStage1VotingMaps_" where the final numeral is associated with the brain
parenchymal labeling given previously.  Similarly, the additional RF prior segmentation
feature probability maps are labeled as "_RFBrainSegmentationPosteriors_".
The Stage 2 feature importance plot follows similar
trends as that for Stage 1 with the T1 images not contributing much to the identification
of white matter hyperintensity voxels.  The initial voting maps from Stage 1 are extremely
important with the top 3 being the estimated locations of the 1) gray matter, 2) white matter, and
3) white matter hyperintensities.  Since these tissue type can be conflated based on intensity
alone it is intuitive that such features would be important.

@@

# Conclusions

### 2016-06-30 figureCaptions.Rmd
ENH:  More revisions.

@@

__Figure 5:__ Average _MeanDecreaseAccuracy_ plots generated from the creation of all 24 random
forest models for Stage 1 during the leave-one-out evaluation.  These plots
are useful in providing a quantitative assessment of the predictive importance of each feature.
Features are ranked in descending order of importance.
The horizontal error bars provide the $95^{th}$ percentile
and illustrate the
stability of the feature importance across the leave-one-out models.
At this initial stage only 31 feature images are used.

__Figure 6:__  Average _MeanDecreaseAccuracy_ plots generated from the creation of all 24 random
forest models for Stage 2 during the leave-one-out evaluation.  These plots
are useful in providing a quantitative assessment of the predictive importance of each feature.
Features are ranked in descending order of importance.
The horizontal error bars provide the $95^{th}$ percentile
 and illustrate the
stability of the feature importance across the leave-one-out models.
We augment the 31 feature images from the first stage by adding an additional
seven voting maps and 7 segmentation posteriors from application of the Bayesian-based
segmentation for a total of 45 images for the second stage.

__Figure 7:__ (a) FLAIR image slice illustrating WMHs which have been manually delineated.
The region around the WMHs is enlarged (b) in the original FLAIR and the
(c) contralateral FLAIR difference image.

__Figure 8:__ Voxelwise comparison with manual delineation of white matter hyperintensities.
On the left are the
calculated Dice values over all white matter hyperintensities.  Note the improvement
in the Dice metric from the employment of the Stage 2 component of the processing pipeline.
(Right) Similar results can be seen by comparing the total lesion load volume between manual
and automated detection strategies.  Although some outliers are found after the Stage 2
processing in a couple subjects, the number of outliers caused by false positives is
decreased significantly with the second stage processing.

>>>>

---
output:
  word_document:
    fig_caption: true
  pdf_document:
    fig_caption: true
    latex_engine: xelatex
    keep_tex: yes
header-includes:
   - \usepackage{booktabs}
   - \usepackage[final]{changes}
   - \usepackage[font={small},labelfont=bf,labelsep=colon]{caption}
   - \linespread{2}
   - \usepackage[compact]{titlesec}
   - \usepackage{enumitem}
   - \usepackage{tikz}
   - \def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;}
   - \setlist{nolistsep}
   - \titlespacing{\section}{2pt}{*0}{*0}
   - \titlespacing{\subsection}{2pt}{*0}{*0}
   - \titlespacing{\subsubsection}{2pt}{*0}{*0}
   - \setlength{\parskip}{3pt}
   - \setremarkmarkup{(#2)}
bibliography: references.bib
csl: council-of-science-editors.csl
fontsize: 12pt
mainfont: Georgia
---

<style type="text/css">
body,
code.bash{
  font-size: 8px;
}
pre {
  font-size: 8px
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set( cache=TRUE )
```

@@

__Figure 5:__ \textcolor{blue}{Evaluation measures for both Stages of the leave-one-out protocol of the described protocol in the Methods
section:  (a) sensitivity, (b) positive predictive value, (c) $F_1$ score, and (d) relative
volume difference.   These quantitative assessments are given for three quantile ranges
spanning the range of the manually-derived lesion volumes.    Overall improvement in all
three whole lesion-based measuers is seen as the
second Stage RF model is applied for all three quantile ranges.
The relative volume difference corresponding to the Stage 2 results tend to predict a
decreased predicted volume over the Stage 1 results.}

__Figure 6:__ Average _MeanDecreaseAccuracy_ plots generated from the creation of all 24 random
forest models for Stage 1 during the leave-one-out evaluation.  These plots
are useful in providing a quantitative assessment of the predictive importance of each feature.
Features are ranked in descending order of importance.
The horizontal error bars provide the $95^{th}$ percentile
and illustrate the
stability of the feature importance across the leave-one-out models.
At this initial stage only 31 feature images are used.

__Figure 7:__  Average _MeanDecreaseAccuracy_ plots generated from the creation of all 24 random
forest models for Stage 2 during the leave-one-out evaluation.  These plots
are useful in providing a quantitative assessment of the predictive importance of each feature.
Features are ranked in descending order of importance.
The horizontal error bars provide the $95^{th}$ percentile
 and illustrate the
stability of the feature importance across the leave-one-out models.
We augment the 31 feature images from the first stage by adding an additional
seven voting maps and 7 segmentation posteriors from application of the Bayesian-based
segmentation for a total of 45 images for the second stage.

__Figure 8:__ (a) FLAIR image slice illustrating WMHs which have been manually delineated.
The region around the WMHs is enlarged (b) in the original FLAIR and the
(c) contralateral FLAIR difference image.
